---
title: "Data resources"
output: 
  github_document: 
    toc: yes
bibliography: data.bib
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


```{r}
library(tidyverse)
```

# Introduction

This repo contains explorations of data that is published openly. Topics include: Waste Management, Air Quality, Wastewater Management, Faecal Sludge Management, WASH (Water, Sanitation, and Hygiene).

For now, all exploration is done in a single R Markdown script rendered to this README file. As the document grows, it will be split into sets of scripts per topic.

# Waste Management

## Abidjan, Cote d'Ivoire, waste characterization data, Feb-Mar 2018

Reference: [@innovations2020abidjan]

Results of municipal solid waste characterization study conducted in Abidjan, Cote d'Ivoire, in February-March, 2018 by [Waste2Worth Innovations](https://www.w2wglobal.com/).


```{r}

dat <- read_csv("data/raw/@innovations2020abidjan/Abidjan, Cote d'Ivoire, waste characterization data, Feb-Mar 2018.csv") %>% 
  janitor::clean_names() %>% 
  group_by(sample) %>% 
  mutate(
    percent = mass_kg / sum(mass_kg) * 100
  ) %>% 
  ungroup()

```

### Exploration

- 2080 observations
- 8 days of data collection
- 80 samples with measurements in 26 waste categories each

```{r}
dat %>% 
  count(date)

dat %>% 
  count(sample)

dat %>% 
  count(date, sample)

dat %>% 
  count(category) 

dat %>% 
  arrange(sample)

```

### Summary tables

```{r}

dat %>% 
  group_by(sample) %>% 
  summarise(
    sum_mass_kg = sum(mass_kg)
  ) %>% 
  arrange(desc(sum_mass_kg)) 
  
dat %>% 
  arrange(sample) %>% 
  group_by(category) %>% 
  summarise(
    count = n(),
    mean = mean(percent),
    sd = sd(percent),
    min = min(percent),
    max = max(percent)
  ) %>% 
  arrange(desc(mean)) %>% 
  knitr::kable(digits = 1)

```

### Data visualisation

```{r}

dat %>% 
  ggplot(aes(x = sample, y = percent, fill = category)) +
  geom_col()

dat %>% 
  ggplot(aes(x = reorder(category, percent), y = percent)) +
  coord_flip() +
  geom_jitter(width = 0.1, alpha = 0.3)




```

## @ramadan2022activity

Reference: [@ramadan2022activity]

```{r}

readxl::read_excel("data/raw/@ramadan2022activity/10163_2022_1371_MOESM2_ESM.xls")

```


## @stadtzurich2022abfallgefasse



```{r}
dat_zurich <- read_csv(file = "data/raw/@stadtzurich2022abfallgefasse/71dc10de-a5e2-11ec-94cb-005056b0ce82/data/erz.abfallgefaess_p.csv")

```

```{r}

dat_zurich
```


## What a waste!

What a Waste is a global project to aggregate data on solid waste management from around the world. This database features the statistics collected through the effort, covering nearly all countries and over 330 cities. Beyond waste generation, there are a long list of additina indicators in this dataset.

-   Source: https://datacatalog.worldbank.org/search/dataset/0039597

```{r}

# This code is copied from application exercise week 10

data_country <- read_csv("data/raw_data/country_level_data_0.csv")

data_city <-  read_csv("data/raw_data/city_level_data_0_0.csv")

codebook_country <- read_csv("data/raw_data/country_level_codebook.csv")

codebook_city <- read_csv("data/raw_data/city_level_codebook_0.csv")

```

Goal: Select data about waste composition for further analysis. 

```{r}

data_city %>% 
  select(country_name, income_id, city_name)

data_city %>% 
  select(-iso3c, -region_id)

data_country %>% 
  select(country_name, ends_with("percent"))

data_country %>% 
  select(country_name, income_id, starts_with("waste_treatment")) %>% 
  pivot_longer(cols = !country_name:income_id,
               names_to = "waste_treatment",
               values_to = "percent") %>% 
  mutate(waste_treatment = str_remove(string = waste_treatment, pattern = "waste_treatment_")) %>% 
  mutate(waste_treatment = str_remove(string = waste_treatment, pattern = "_percent")) %>% 
  
  group_by(waste_treatment) %>% 
  summarise(
    count = n(),
    percent = mean(percent, na.rm = TRUE)
  ) %>% 
  
  ggplot(aes(x = income_id, y = percent, fill = waste_treatment)) +
  geom_col()






data_city %>% 
  names()

data_city %>% 
  select(country_name:city_name, 
         waste_treatment_advanced_thermal_treatment_percent:waste_treatment_waterways_marine_percent) 

```


# References
